### 2013-09-19 main.tex
Polish intro, add blank chapter on evaluation
new title, polish conclusion

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction for Semantic Image Segmentation}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

@@

The first part introduces a clustering algorithm that optimizes a novel
information-theoretic objective function. We show that the proposed algorithm
has clear advantages over standard algorithm from the literature on a wide
array of datasets. Clustering algorithms are an important building block for
higher level computer vision applications, in particular for semantic
segmentation.

The second part of this work proposes an algorithm for automatic segmentation
and recognition of object classes in natural images, that learns a segmentation
model solely from annotation in form of presence and absence of object classes
in images. To the best of our knowledge, this is the first algorithm that is
also able to addresses complex object classes.

The third and main part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.

We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.
The introduced methods provide results advancing the state-of-the art on two
complex benchmark datasets for semantic segmention, the MSRC-21 Dataset of RGB
images and the NYU V2 Dataset or RGB-D images of indoor scenes.

Finally, we introduce a software library that allows us to perform extensive empirical
comparisons of state-of-the-art structured learning approaches. This allows us
to characterize their practical properties in a range of applications, in
particular for semantic segmentation and object class segmentation.

@@

@@

\chapter{Conclusion and Future Direction}
Current research in computer vision progresses at an incredible speed, with new
methods and applications emerging constantly.  Still, there are large
differences in the paradigms employed by different researchers.  While it is
generally acknowledged that context is one of the most important cues in
recognizing objects and interpreting scenes, there is little consensus on how
contextual information should be modelled, or even learned.
The structured prediction paradigm allows a principled approach to
incorporating predictions within an image, and to include context information.
We investigated the potential of structural methods for the application of
semantic image segmentation and object recognition, applications to which
structured methods have been applied successfully before.  We focus on
\emph{learning} of structural models and the interaction of inference and
learning in the neighborhood models typically employed for semantic
segmentation.
The thesis presents a software implementation of a variety of popular learning
algorithms for structural support vector machines, together with a thorough
evaluation of their properties in the presence of approximate inference.
We demonstrate that exact learning is possible, even in the presence of loops in the
underlying factor graph. We also demonstrate the power of conditioning interactions
on data, learning spacial interactions in a RGB-D setting.
%TODO STATE OF THE F*ING ART!

There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results, in particular in
extending the second and third part:
\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
google
\paragraph{Inference Machines}
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph{Non-Linear Models}
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to solve even object-centric tasks such as object class
segmentation with DTFs.

\paragraph{Latent Variable Models}
Something latent
\paragraph{Feature Design}
features are important
which low-level cues?
superpixels?

>>>>

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

@@

The first part introduces a clustering algorithm that optimizes a novel
information-theoretic objective function. We show that the proposed algorithm
has clear advantages over standard algorithm from the literature on a wide
array of datasets. Clustering algorithms are an important building block for
higher-level computer vision applications, in particular for semantic
segmentation.

The second part of this work proposes an algorithm for automatic segmentation
and recognition of object classes in natural images, that learns a segmentation
model solely from annotation in form of presence and absence of object classes
in images.

The third and main part of this work investigates one of the most popular
approaches to the task of object class segmentation and semantic segmentation,
based on conditional random fields and structured prediction.
%
We investigate several learning algorithms, in particular in combination with
approximate inference procedures. We show how structured models for image
segmentation can be learned exactly in practical settings, even in the presence
of many loops in the underlying neighborhood graphs.
The introduced methods provide results advancing the state-of-the art on two
complex benchmark datasets for semantic segmentation, the MSRC-21 Dataset of RGB
images and the NYU V2 Dataset or RGB-D images of indoor scenes.
%
Finally, we introduce a software library that allows us to perform extensive empirical
comparisons of state-of-the-art structured learning approaches. This allows us
to characterize their practical properties in a range of applications, in
particular for semantic segmentation and object class segmentation.

@@

\include{evaluation/paper}

@@

\chapter{Conclusion and Future Direction}
\section{Summary}
This thesis explored the use of structured prediction methods for semantic
segmentation and object class segmentation of natural images, as an important
step towards general scene understanding.
We use the paradigm of structured prediction, which allows for a principled
integration of context and object relations.
%Current research in computer vision progresses at an incredible speed, with new
%methods and applications emerging constantly. Still, there are large
%differences in the paradigms employed by different researchers. While it is
%generally acknowledged that context is one of the most important cues in
%recognizing objects and interpreting scenes, there is little consensus on how
%contextual information should be modelled, or even learned.
%The structured prediction paradigm allows a principled approach to
%incorporating predictions within an image, and to include context information.
%We investigated the potential of structural methods for the application of
%semantic image segmentation and object recognition, applications to which
%structured methods have been applied successfully before.
We focus on \emph{learning} of structural models and the interaction of
inference and learning in the neighborhood models typically employed for
semantic segmentation.
We presents an open source software implementation of a variety of
popular learning algorithms for structural support vector machines, together
with a thorough evaluation of their properties in the presence of approximate
inference. Our software provides a solid basis for future research into learning,
inference and models for computer vision, by providing extensive examples and
benchmarks.

We demonstrate that effective use of available inference mechanisms enables
exact learning, even in the presence of loops in the underlying
factor graph, leading to state-of-the art results on the MSRC-21 dataset.
We also demonstrate the power of conditional interactions, learning spacial
interactions in a RGB-D setting. Our approach improves upon the state-of-the
art on the NYU V2 benchmark for annotation of semantic structure classes.

We also presented a novel approach for clustering based on information theoretic
principles. Our algorithm improves upon methods from the literature on a wide
range of datasets. This indicates that also in the task of extracting
superpixels, we can hope to achieve better results than the $k$-means based
SLIC algorithm that we used.

As manual annotation of images for learning semantic segmentation and object
class recognition is laborious and error-prone, we suggested a method to
learn object class segmentation for complex object classes from image-level
annotations alone.

\section{Future Directions}
There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results:
\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
We demonstrated a new method for object class segmentation using only weak
supervision. One of the main advantages of such a method is that it is
potentially able to exploit the large amount of weakly labeled data that
is available on the internet. Using additional training data, and evaluating on
the given, manually annotated data, would be an interesting topic for future
research.
\paragraph{Inference Machines}
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph{Non-Linear Models}
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels. Allowing non-linear
interactions increases the representational power of a CRF, possibly leading
to better segmentation results.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher-level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to address even object-centric tasks such as object class
segmentation with DTFs.

\paragraph{Higher Order Potentials and Latent Variable Models}
While non-linear potentials would allow for more complex interactions between
inputs and labelings, introducing higher order potentials or latent variables
allow the model to express more complex interactions within the output
variables. Possible examples are consistency of larger regions, learning parts
or learning of scene classes and cooccurcences.
In principle, higher order potentials and latent variable models are
equivalent, in that each energy function expressed in either form can be
transformed into an energy function of the other kind. In practice, learning of
higher order potentials for semantic segmentation has received little
attention, while most approaches using latent variables are often limited by
the non-convexity of learning.  It would be interesting to compare current
methods using latent variable and higher order approaches, and see how these
interact with different inference and learning schemes.

\paragraph{Feature Design}
This work mostly focused on learning methods, and less on the input, with
the exception of Chapter~\ref{ch:nyu}, that explores the use of 3D features
for semantic segmentation of indoor scenes. It is clear, however, that
the input feature play an important role in the performance of any system.
Using our approach for exact learning of loopy graphs, it seems to be worthwhile to
revisit the works of \citet{nowozin2010parameter} and \citet{lucchi2011spatial},
that evaluate the impact of input features and piecewise training, and of the importance
of global constraints versus global features, respectively.
In particular the importance of features for pairwise potentials has been
somewhat overlooked, often being reduced to a single constant or contrast
sensitive feature.
